---
title: "R Notebook"
output:
  pdf_document: default
  word_document: default
  html_notebook: default
---


**Author:** Andres Felipe Alba Hernández  
**Department:** Electrical Engineering  
**Date:** October 1st, 2018  
**Course:** ISYE670 Data Science for Engineers  
**Professor:** Dr. Christine Nguyen  
**Northern Illinois University**  

```{r}
rm(list = ls())
library(dplyr,lib.loc="/home/leasanspy/DataScience_NIU/Rpackages") #Loading the library
```


1) (2 points) Import the data with the headers. How is the data organized? (i.e. what are the
column names? What does each row of the data represent?)  

From the information below I infered that the data is organized in a way that every observation collected an index which reference the first letter of the state, the state, and the total income by year.

```{r}
h_data <- read.csv("historical_data.csv") 
names(h_data)
str(h_data)
summary(h_data)
```

2) Lookup the sample_n and sample_frac functions.
a) (2 points) Use the appropriate function to randomly choose 15 rows.
b) (2 points) Use the appropriate function to randomly choose 40%

```{r}
Sn <- sample_n(h_data,15) #a
Sf <- sample_frac(h_data,0.4) #b
str(Sn) #a
str(Sf) #b
```

 3) Output certain columns.
a) (2 points) Output only the values for the year 2005 and States
b) (2 points) Output only the values for all the years. (Hint, use the - )
c) (2 points) Output only the values for columns that begin with a Y


```{r}
#Answer a)
select(h_data,State,Y2005)
```

```{r}
#Answer b
select(h_data,-Index,-State)
```

```{r}
#Answer c
select(h_data,starts_with("Y"))
```

4) Output only certain observations.
a) (2 points) Output only the observations with an index values of A and C and N.
b) (2 points) Output only the observations related to Illinois and California.

```{r}
#Answer a
filter(h_data, Index=="A"|Index=="C"|Index=="N")
#Answer b
filter(h_data, State=="Illinois"|State=="California")
```

5) (2 points) Output data that shows only the index, state, and 2010 income values greater than $1,500,000.
a) (Extra Credit: 2 points) The same code can be re-written concisely using the pipe operator
%>%. What is that command?

```{r}
select(filter(h_data, Y2010>=1500000),Index,State,Y2010) #normal way
select(h_data,Index,State,Y2010) %>% filter(Y2010>=1500000) #with the pipe opearator
```

6) (3 points) Output only the state and 2006 values, and arrange the values in descending order of 2006 values. What are the 3 highest income generating states?  

The states of Minnesorta, Pennsylvania, and West Virginia have the highest income:

```{r}
select(h_data,State,Y2006) %>% arrange(desc(Y2006))
```

7) (3 points) Using the mutate function, create a new column “ratio” which divides the income of 2015 by 2014. Use the appropriate combination of commands so only the state, 2014, 2015,
and the new column is saved or output.


```{r}
mutate(select(h_data,Y2014,Y2015),ratio=Y2015/Y2014)
```

8) (3 points) Using group_by and summarise, create a table grouped by Index and then
summarizes each group by taking the mean of Y2007

```{r}
h_data %>% group_by(Index) %>% summarise(mean_2007=mean(Y2007))
```

9) (3 points) Group the observations by index. Output data that shows the average ratio of 2015 divided by 2014 for each group.

```{r}
h_data %>% group_by(Index) %>% mutate(ratio=Y2015/Y2014) #This is the answer.
```

```{r}
#This is to show more clearly the ratio. 
h_data %>% group_by(Index) %>% mutate(ratio=Y2015/Y2014) %>% select(ratio)
```


