---
title: "Homework 12 k-means clusters"
output:
  word_document: default
  html_document:
    df_print: paged
  pdf_document: default
always_allow_html: yes
---
**Author:** Andres Felipe Alba Hernández  
**Department:** Electrical Engineering  
**Date:** November 25th, 2018  
**Course:** ISYE670 Data Science for Engineers  
**Professor:** Dr. Christine Nguyen  
**Northern Illinois University**


**POINT A**  

```{r}
rm(list=ls()) 
```

```{r}
ds=read.csv2("Wholesale_customers_data.csv",sep=",",header = TRUE)
```

```{r}
str(ds)
#ds
```
```{r}
n_ds=ds[,c(3:8)]
str(n_ds) #creating new data set
```

```{r}
set.seed(250) #Creating the seed for whatever random task in the code
```

**POINT B**  
The data was from the website:  
i. (2 points) What country does the information come from?  
The information is from Portugal.  
ii. (2 points) Look at the attribute information. What do the attributes “Fresh” and
“Region” mean?  

*Fresh: annual spending (m.u.) on fresh products.  
*Region: customers Region region.  Lisbon, Oporto or Other Region.  


**POINT C**  

Perform the k-means clustering using the “kmeans” command. Do not use the NbClust
command. Let k = 3 and answer the following questions.  

```{r}
#Scaling the data
s_ds<-scale(n_ds)
```


```{r}
#perform the clustering
ds_fit=kmeans(x=s_ds,centers = 3,nstart = 40)
```

```{r}
ds_fit
```

i. (3 points) How many data points are in each cluster?   
After performing the clustering 8 points belong to the first cluster, 345 points belong to the second cluster, and 87 points belong to the third cluster.  

ii. (2 points) What cluster is observation 80 in?  
The observation 80 belongs to the cluster 2.  

```{r}
ds_fit$cluster[80]
```

iii. (3 points) What are the centers of the clusters?  

The centers are the vectors 1,2,3 describe below.  
```{r}
ds_fit$centers
```

iv. (8 points) Compare the results of your kmeans algorithm results with the Region column of the original data. What is the composition of each cluster with respect to the Region type.  

```{r}
sum(ds$Region==1)
sum(ds$Region==2)
sum(ds$Region==3)
```

```{r}
table(ds_fit$cluster,ds$Region)
```

The rows of the table above reference to the cluster while the columns reference the Region. For example, the cluster one have 0 elements of region 1, two elements of region 2 and 7 elements of region 3.  

**POINT D**  

(10 points) Perform the k-means clustering using the “kmeans” command. Do not use the NbClust command. Perform the algorithm for k values 1 through 10. Using within sum of squares, between sum of squares and total sum of squares, determine which value of k you would recommend. Explain why.  


```{r}
for (i in 1:10)
{
  test<-kmeans(x=s_ds,centers = i,nstart = 40)
  print(i)
  #print(test$withinss)
  print(test$tot.withinss) #Total Within cluster sum of squares. WSS
  print(test$betweenss) #Between sum of squares
  print(test$totss) #Total sum of squares
  print("----------------------")
}
```

The data for each k was printed above. Small WSS indicates more compact and dense clusters so better ones for classification. Use Within Sum of Squares: measures the variability of the observations within a clusterThe lowest value was obtained for k=10. Because the total sum squares is the same then the between-cluster sum of squares have the highest value for k = 10  







