---
title: "Homework9 KNN"
output:
  word_document: default
  html_document:
    df_print: paged
  pdf_document: default
always_allow_html: yes
---
**Author:** Andres Felipe Alba Hernández  
**Department:** Electrical Engineering  
**Date:** October 26th, 2018  
**Course:** ISYE670 Data Science for Engineers  
**Professor:** Dr. Christine Nguyen  
**Northern Illinois University**  

```{r}
#Initializing
rm(list=ls()) #deleting environment variables
#Installing the library (only once)
#install.packages("FNN") 
library(FNN)
library(dplyr) #Loading the library
```

**Data Description**

a) Did the data come with a header or not? The does not have headers, then the headers may be added manually as its done below.  


```{r}
nombres <- c("Code_number","Clump Thickness", "uniformity cell size","uniformity cell shape", "Marginal Adhesion", "Epithelial Cell Size", "Bare Nuclei", "Bland Chromatin", "Normal Nuceloili", "Mitoses", "Class Cell")
cancer_data <- read.csv("breast-cancer-wisconsin.data")
names(cancer_data)<- nombres
str(cancer_data)
```

b) What is the target variable? How well is the data distributed among the target variable?

The target value is Class Cell where the value 2 indicate benign cells and the value 4 indicate malignant cells. The features can help to classify the sample, for example in the summary of the data below it is clear that there are differences in the distribution depending on the class, for example for the malignant class the mean of the Clump thickness is 7.1 while for the benignant it is 2.9. 

```{r}
summary(cancer_data)
Malignant<-cancer_data %>% filter(`Class Cell`== 4)
Bening <- cancer_data %>% filter(`Class Cell`== 2)
```
```{r}
print('Malignant -------')
summary(Malignant)
print('Benign -------')
summary(Bening)
```



c) What are the possible attributes? 

```{r}
str(cancer_data)
```

**Note:** I extracted the Attributed table below from:
https://archive.ics.uci.edu/ml/machine-learning-databases/breast-cancer-wisconsin/breast-cancer-wisconsin.names

As it can be appreciated below each data value take a value between 1 and 10. However the Bare Nuclei is the only factor value. The classes from 1-6 can contain missing values that are marked as "?"" according to the documentation, however, this can be fixed changing this data for a value such as zero. 


Attribute Information: (class attribute has been moved to last column)

   #  Attribute                     Domain
   -- -----------------------------------------
   1. Sample code number            id number
   2. Clump Thickness               1 - 10
   3. Uniformity of Cell Size       1 - 10
   4. Uniformity of Cell Shape      1 - 10
   5. Marginal Adhesion             1 - 10
   6. Single Epithelial Cell Size   1 - 10
   7. Bare Nuclei                   1 - 10
   8. Bland Chromatin               1 - 10
   9. Normal Nucleoli               1 - 10
  10. Mitoses                       1 - 10
  11. Class:                        (2 for benign, 4 for malignant)

**Data preparation**

In order to use the KNN algorithm the dataset should not contain NA values and the data needs to be numeric. Therefore some preprocessing has to be done in that direction.  

```{r}
length(names(cancer_data)) #698
#names(cancer_data)
cancer_data[cancer_data=="?"]<-0 #This will create NA levels when you have factor levels and you do not have 0 in the possible levels. 
cancer_data <- na.omit(cancer_data) #I would like to omit the NA values. 
```

```{r}
cancer_data$`Bare Nuclei` <- as.numeric(as.character(cancer_data$`Bare Nuclei`))
#as.numeric(cancer_data$`Bare Nuclei`) #This will give you the index level instead of the value,( that is the reason why you may use as.character() firts
```

**Model**

a) Create a training set and test set using a 75%/25% split, respectively

The following code split the data in 75 and 25%. Then it store the training data, the labels (classification) of the training, the Validation data, and the labels of the validation data in different data frames. 

```{r}
set.seed(1234)
index <- sample(2, nrow(cancer_data), replace = TRUE, prob = c(0.75, 0.25))

#Index 1 refer to the 75% of the data
Cancer_TrainSet <- cancer_data[index == 1, 1:length(names(cancer_data))-1] #This skip the last column, label
Cancer_TrainLabel <- cancer_data[index == 1, length(names(cancer_data))] #it takes only the label

#Index 2 refer to the 25% of the data- 
Cancer_ValidationSet <- cancer_data[index == 2, 1:length(names(cancer_data))-1]
Cancer_ValidationLabel <- cancer_data[index == 2, length(names(cancer_data))]
```

It is good to take a look on the resulting data to be sure that the preprocessing works correctly.

```{r}
str(Cancer_TrainLabel)
str(Cancer_TrainSet)
Missing <- cancer_data %>% filter(`Bare Nuclei`== "?")
str(Missing)
```

```{r}
sum(Cancer_TrainSet=="?")
sum(Cancer_ValidationSet=="?")
#summary(is.na(Cancer_TrainSet))
```

```{r}
sum(is.na(Cancer_TrainSet))
sum(is.na(Cancer_ValidationSet))
```


b) Perform kNN. Consider k values from 1 to 50. Find the accuracy of your model for these
values and plot this as a line graph. 

```{r}
#Inizilizing Values
size_validation=length(Cancer_ValidationLabel) #number of elements in the validation set
classification_mistake<- 0 #inizializing the number of mistakes in zero
accuracy_prediction <- 0 #inizializing the accuracy of prediction in zero
Accuracy_array<-c() #inizializing the array for store the k accuracy 

#Iterating among differe
for(i in 1:50)
  {
  #Evaluating the model for k number of neigbors. 
  predict_i<-knn(Cancer_TrainSet, Cancer_ValidationSet, cl = Cancer_TrainLabel, k = i, prob = FALSE)
  classification_mistake<-length((which(as.numeric(as.character(predict_i)) != Cancer_ValidationLabel))) 
  accuracy_prediction <-classification_mistake/size_validation
  #print(accuracy_prediction)
  Accuracy_array[i]<-accuracy_prediction
  }
print("done with the prediction array")
```


```{r}
#Building the plot
plot(Accuracy_array)
```


c) Which k value provides the best accuracy? What is the accuracy value?

The best accuracy was found at k=7 with an accuracy of 42%. 

```{r}
#print(Accuracy_array)
max(Accuracy_array)
which.max(Accuracy_array)
```


d) Is the model good or bad? Why do you think so and what would you do to improve the
model?

The model is bad because it fails in more than half of the predictions. I explore moving k from 1 to 502 and still get the same accuracy, after create the histogram (below) of the training labels I see that the distribution between the two clases is not completely balance. Maybe having a more balance Training dataset may increase the performance. After ploting the histogram of the mistake labels for k=7 (Best performance) I see that the algorithm missclassify more Malignant than Bening samples. In general you have more Bening samples, so it is more likely to find neigborhs of this class. 

```{r}
size_validation=length(Cancer_ValidationLabel) #number of elements in the validation set
classification_mistake<- 0 #inizializing the number of mistakes in zero
accuracy_prediction <- 0 #inizializing the accuracy of prediction in zero
Accuracy_array<-c() #inizializing the array for store the k accuracy 

for(i in 1:502)
  {
  #Evaluating the model for k number of neigbors. 
  predict_i<-knn(Cancer_TrainSet, Cancer_ValidationSet, cl = Cancer_TrainLabel, k = i, prob = FALSE)
  classification_mistake<-length((which(as.numeric(as.character(predict_i)) != Cancer_ValidationLabel))) 
  accuracy_prediction <-classification_mistake/size_validation
  #print(accuracy_prediction)
  Accuracy_array[i]<-accuracy_prediction
  }
print("done with the prediction array")

#print(Accuracy_array)
max(Accuracy_array)
which.max(Accuracy_array)
```

```{r}
#Building the plot
plot(Accuracy_array)
```

```{r}
hist(Cancer_TrainLabel)
```

```{r}
  #Evaluating the model for k=7 number of neigbors. 
  i=7
  predict_i<-knn(Cancer_TrainSet, Cancer_ValidationSet, cl = Cancer_TrainLabel, k = i, prob = FALSE)
  mistake_data<-(which(as.numeric(as.character(predict_i)) != Cancer_ValidationLabel)) 
  classification_mistake<-length((which(as.numeric(as.character(predict_i)) != Cancer_ValidationLabel))) 
  accuracy_prediction <-classification_mistake/size_validation
  #print(accuracy_prediction)
  Accuracy_array[i]<-accuracy_prediction
```

```{r}
#Remember that 4 is Malignant
hist(Cancer_ValidationLabel[mistake_data])
```

```{r}
#Malignant 239
#Benign 443
#I will split the dataset into Malignant and bening. 
Malignant<-cancer_data %>% filter(`Class Cell`== 4)
Bening <- cancer_data %>% filter(`Class Cell`== 2)
#lets take same amount of samples from each one to create the training dataset
#index <- sample(2, nrow(cancer_data), replace = TRUE, prob = c(0.75, 0.25))
index1 <- sample(2,nrow(Malignant),replace = TRUE,prob=c(0.5,0.5))
index2 <- sample(2,nrow(Bening),replace = TRUE,prob=c(0.27,0.73))

Malignant_TrainSet <- Malignant[index1 == 1, 1:length(names(cancer_data))-1] #This skip the last column, label
Malignant_TrainLabel <- Malignant[index1 == 1, length(names(cancer_data))] #it takes only the label
Malignant_ValidationSet <- Malignant[index1 == 2, 1:length(names(cancer_data))-1]
Malignant_ValidationLabel <- Malignant[index1 == 2, length(names(cancer_data))]

Bening_TrainSet <- Bening[index2 == 1, 1:length(names(cancer_data))-1] #This skip the last column, label
Bening_TrainLabel <- Bening[index2 == 1, length(names(cancer_data))] #it takes only the label
Bening_ValidationSet <- Bening[index2 == 2, 1:length(names(cancer_data))-1]
Bening_ValidationLabel <- Bening[index2 == 2, length(names(cancer_data))]

Training_set<-rbind(Malignant_TrainSet,Bening_TrainSet)
Validation_set<-rbind(Malignant_ValidationSet,Bening_ValidationSet)

#Training_set_Label<-rbind(Malignant_TrainLabel,Bening_TrainLabel)
#Validation_set_Label<-rbind(Malignant_ValidationLabel,Bening_ValidationLabel)

Training_set_Label<-c(Malignant_TrainLabel,Bening_TrainLabel)
Validation_set_Label<- c(Malignant_ValidationLabel,Bening_ValidationLabel)
```

```{r}
#str(Training_set)
#str(Training_set_Label)
#Training_set_Label
#Training_set
#Validation_set_Label
str(Cancer_TrainLabel)
str(Training_set_Label)
str(Validation_set)
```

```{r}
  #Evaluating the model for k=7 number of neigbors. 
#i<-7
classification_mistake=0
accuracy_prediction_modif<-c()

for(i in 1:length(Training_set_Label))
  {
    size_validation<-length(Validation_set_Label)
    #predict_i<-knn(Cancer_TrainSet, Cancer_ValidationSet, cl = Cancer_TrainLabel, k = i, prob = FALSE)
    predict_i<-knn(Training_set, Validation_set, cl = Training_set_Label, k = i, prob = FALSE)
    #mistake_data<-(which(as.numeric(as.character(predict_i)) != Validation_set_Label)) 
    classification_mistake<-length((which(as.numeric(as.character(predict_i)) != Validation_set_Label))) 
    accuracy_prediction_modif[i] <-classification_mistake/size_validation
    #print(classification_mistake)
    #print(size_validation)
    #print(accuracy_prediction_modif)
    #accuracy_prediction_modif
}

max(accuracy_prediction_modif)
which.max(accuracy_prediction_modif)
```

As could be observed above, changing the training data set for a more balance improve the performance from 42% to 73%, the max accuracy was for k=175 in this case. I will conclude that this change improve the performance. 

