---
title: "ICA4 Module 4"
output:
  word_document: default
  html_notebook: default
  pdf_document: default
---

  word_document: default
  pdf_document: default
---

**Author:** Andres Felipe Alba Hernández  
**Department:** Electrical Engineering  
**Date:** September 23, 2018  
**Course:** ISYE670 Data Science for Engineers  
**Professor:** Dr. Christine Nguyen  
**Northern Illinois University**  

```{r}
#Initializing
rm(list=ls()) #Deleting all variables in the enviroment
#install.packages("tidyr", lib="/home/leasanspy/DataScience_NIU/Rpackages") #Installing the library
library(tidyr,lib.loc="/home/leasanspy/DataScience_NIU/Rpackages") #Calling the library 
```



1. Using R functions, what are the header names?  

The names are the ones listed below: 
```{r}
rev_hist=read.csv("Revenue_Historical_2000.csv", header = TRUE)
names(rev_hist)
```

2. Output a summary of the data structure. What is the mean of the values in Month 2?  

The mean of the month two is: 5558
```{r}
#str(rev_hist)
summary(rev_hist)
```

3. Use the gather function to create a new table with the desire structure:  
```{r}
#gather(data,key: out_name_column, value: out_data_name, list_of_names_with_columns_gathered)
new_table_gather=gather(rev_hist, Month, Revenue,Month.1:Month.12) 
#new_table_gather
```

4. Use the separate function to separate the values under the “Month” column in part 3.  
```{r}
#names(new_table_gather)
#separate( data, col, into, sep = [regular expression], remove = TRUE, convert = FALSE)
new_table_separate=separate(new_table_gather,Month,c("Time","Time_Interval"),remove=TRUE,convert=FALSE)
new_table_separate
```

5. Use the unite function on the data from part 4 to merge “Time” and “Time_Inverval” create a
new column to read as “1_Month”
```{r}
#unite( data, name_of_new_column, names_of_col_merging_[LIST] , sep = “ “, remove = TRUE)
new_table_unite=unite(new_table_separate, "New_Time_Month", c("Time_Interval","Time"), sep = "_",remove=TRUE)
new_table_unite
```

6. Use the spread function on the data from part 5 such that the columns are the years and
keep the month in a column.

```{r}
#spread( data, key: columns_to_spread, value: values_of_spreadColumns, fill= NA, convert = FALSE)
spread(new_table_unite,Year,Revenue)
```