---
title: "ICA 5 Module 5 dplyr_library"
output:
  word_document: default
  html_notebook: default
  pdf_document: default
---

  word_document: default
  pdf_document: default
---

**Author:** Andres Felipe Alba Hernández  
**Department:** Electrical Engineering  
**Date:** September 30, 2018  
**Course:** ISYE670 Data Science for Engineers  
**Professor:** Dr. Christine Nguyen  
**Northern Illinois University**  

Questions 1,2 and 3 can be compacted in the commands line below.

```{r}
#Initializing
rm(list=ls()) #deleting enviroment variables
#Installing the library (only once)
#install.packages("tidyverse", lib="/home/leasanspy/DataScience_NIU/Rpackages") 
#install.packages("dplyr", lib="/home/leasanspy/DataScience_NIU/Rpackages") 
library(dplyr,lib.loc="/home/leasanspy/DataScience_NIU/Rpackages") #Loading the library
```

4) As may be observed from the commands below the income dataset describes Units sales, Revenue and Expenditure by Region, Department, and Year. 

```{r}
income <-read.csv("data_module_5.csv",header = TRUE)
summary(income)
```

5) The only factor variable in the dataset is the Region and it has three levels {A,B,C}. I consider that Departments could be also a factor variable with levels 1 and 2. 

```{r}
str(income)
#income
```

6) Sample command by number and by percentage:

```{r}
S1 <- sample_n(income,5) #five random samples
S2 <- sample_n(income,5) #other five random samples
S3 <- sample_frac(income, 0.02)
S4 <- sample_frac(income, 0.02)
```

```{r}
print(S1)
print(S2)
print(S3)
print(S4)
```

7) The select function is used to select desired variables  
  a) Retrieve only the Department and Year columns  
  b) Retrieve all columns except the Expenditure column
```{r}
select(income,Department,Year) #answer for a
select(income,-Expenditure) #answer for b
```

8) Using the help function, answer the following questions in your own words.
  a) What is the ends_with() function?
  b) What is the contains() function?
  c) What is the matches() function?
  
All this functions are apply to the columns headers, basically it will select the column name that accomplish certain search using teh helpers.  

For example:  

a) ends_with() will pick the headers that ends with certain string that you give. 
b) contains() will pick the headers that its name contain the given string, you can choose if you want to ignore or use the case differences.  
c) matches() this helper help you to pick the right header by using regular expressions.  
There are other helpers that can be obsered using the help for "select." In order to be more clear, I create one example for each select helper below: 
  
```{r}
#as_tibble(income)
summary(income)
summary(select(income, ends_with(match = "ue",ignore.case = TRUE))) #I extract revenue
summary(select(income, contains("expen",ignore.case = TRUE))) #This should extract Expenditure
summary(select(income, matches("Ye*"))) #This should extract year
```

9) The filter function is used to filter rows based on the criteria provided by the user. What
command will filter the rows/observations where the year is 2000 or 2002?

```{r}
filter(income, Year == "2000" | Year == "2002")
```

10) Suppose you want to filter to keep rows where the Year is 2002, and then select the
Department and Unit_Sales columns, and save it all in a new variable.  

The code below execute both approaches, I use summary only to save space at printing 

```{r}
sd <- select(filter(income, Year == "2002"), Department, Unit_Sales)
summary(sd)
sd_pipe <- income %>% filter(Year=="2002")%>% select(Department, Unit_Sales)
summary(sd_pipe)
```

11) The arrange function is used to arrange or re-order rows by a particular column
Let’s reorder using the Department values. The values should be in ascending order.

```{r}
arrange(income, Department)
```

```{r}
arrange( income, desc(Department))
```

12)The mutate function is used to create new variables that are functions of existing variables. Create a new data frame that has the Expenditure and Revenue columns, and also create a new column “profit”, which is Revenue minus Expenditure.

```{r}
m_df <- mutate(select(income, Expenditure, Revenue), profit=Revenue-Expenditure)
m_df_pipe <- income %>% select(Expenditure,Revenue) %>% mutate(profit=Revenue-Expenditure)
m_df
m_df_pipe
```

Both output have similar results. 

13) Let’s use the group_by and summarise functions to calculate the average expenditure by
Region

```{r}
income %>% group_by(Region) %>% summarise(avg_expenditure=mean(Expenditure))
income %>% group_by(Region) %>% summarise(standarDesviation_expenditure=sd(Expenditure))
income %>% group_by(Region) %>% summarise(minimum_expenditure=min(Expenditure))
income %>% group_by(Region) %>% summarise(maximum_expenditure=max(Expenditure))
income %>% group_by(Region) %>% summarise(Median_of_expenditure=median(Expenditure))
```



