---
title: "ICA_1 AndresAlba"
output:
  word_document: default
  pdf_document: default
  html_document: default
---


## R Markdown

<b>Chair Orders Excersice</b>
#1 
#1.a
```{r}
orderNum <- c(1:5)
numUnits <- c(5,7,3,4,6)
color <- c("brown", "red", "red", "blue", "red")
done <- c(TRUE, TRUE, FALSE, FALSE, FALSE)
```

#1.b
```{r}
ChairOrders <- data.frame(orderNum, numUnits, color, done)
```

#1.c
This remove the datastructures but it keep the Chair Orders
```{r}
out_rm <- rm(orderNum, numUnits, color, done)
```

#1.d
the name of the first columb is orderNum
[1] "orderNum" "numUnits" "color"    "done" 
```{r}
names(ChairOrders)
```
#1.e
```{r}
print(ChairOrders)
```

The same frame with modify names
```{r}
orderNum <- c(1:5)
numUnits <- c(5,7,3,4,6)
color <- c("brown", "red", "red", "blue", "red")
done <- c(TRUE, TRUE, FALSE, FALSE, FALSE)
ChairOrders_modif_names <- data.frame(ordenes=orderNum, unidades=numUnits, colores=color, realizado=done)
```
#2
mean_2a <- mean(ChairOrders)
mean_2b <- mean(ChairOrders$numUnits)
> mean_2a
[1] NA
> mean_2b
[1] 5
```{r}
mean_2a <- mean(ChairOrders)
mean_2b <- mean(ChairOrders$numUnits)
```

#3
#3.a
```{r}
sum_numUnits <- sum(ChairOrders$numUnits)
sum_numUnits
```

#3.b
```{r}
print(ChairOrders$numUnits)
```


#3.c
```{r}
ChairOrders$numUnits[2]
```

#3.d
```{r}
vector_red<-ChairOrders$color=="red"
vector_red
```

```{r}
vector_done<-ChairOrders$done=="FALSE"
vector_done
```
#3.e
```{r}
chairs_notDone <- sum(ChairOrders$numUnits[ChairOrders$done==FALSE])
chairs_notDone
```
#3.f
```{r}
chairs_Done <- sum(ChairOrders$numUnits[ChairOrders$done==TRUE])
chairs_Done
```

#3.g
```{r}
chairs_Red <- sum(ChairOrders$numUnits[ChairOrders$color=="red"])
chairs_Brown <- sum(ChairOrders$numUnits[ChairOrders$color=="brown"])
chairs_Red
chairs_Brown
```

#4.a
```{r}
chair_red_notDone <- sum(ChairOrders$numUnits[ChairOrders$color=="red"])
chair_red_notDone
```

#4.b
```{r}
chair_notRed_notDone <- sum(ChairOrders$numUnits[ChairOrders$color!="red" & ChairOrders$done==FALSE])
chair_notRed_notDone
```

#5
```{r}
summary(ChairOrders)
```


The command will show the min, the first quartile, median, mean, third quartile and maximum for each column vector each vector is like a characteristic of the data so you can have a general idea of your data. 

#6.a
```{r}
help(mode)
```

Description

Get or set the type or storage mode of an object.

Usage

mode(x)
mode(x) <- value
storage.mode(x)
storage.mode(x) <- value
Arguments
.
.
.

#6.b
With the mode function you can obtain the type of element of the object in the data structure. 
```{r}
mode(ChairOrders$color)
```

```{r}
mode(ChairOrders$done)
```

#General Questions
#For saving space I decided to not print the outputs in this session. 
#1
```{r}
mydata <- data.frame(a = numeric(0), b = logical(0), c = character(0))
data_a <- c(12,56,30,18)
data_b <- c(TRUE,TRUE,FALSE,TRUE)
data_c <- c("YES","NO","NO","YES")
my_newdata <- data.frame(data_a,data_b,data_c)
mydata <- edit(my_newdata)
```

#2
```{r}
vector_int=c(8:17)
length(vector_int)
```

#3
```{r}
lt <- "My homework list"
movies <- c("Inglorious Bastards","The life is beautiful","Match Point")
expences <- c(1500,2500,900,1100,1150)
family <- "Vanessa"
my_homework_list <- list(title=lt,movies=movies,expenditures=expences,family_members=family)
```

#4
```{r}
vector_inc=seq(100,110,0.5)
```

