---
title: "IMBD Data Science Project"
output:
  word_document: default
  html_document:
    df_print: paged
  pdf_document: default
always_allow_html: yes
---
**Author:** Andres Felipe Alba Hernández  
**Department:** Electrical Engineering  
**Date:** December 10th, 2018  
**Course:** ISYE670 Data Science for Engineers  
**Professor:** Dr. Christine Nguyen  
**Northern Illinois University**


```{r}
rm(list=ls()) 
```
```{r}
library(knitr)
knitr::opts_chunk$set(echo = TRUE, rows.print=3)
```
```{r}
#Libraries
library(ggplot2)
library(data.table) #For extended data frames 
library(dplyr)
```
#Data Description and preparation

##Data Exploration
```{r}
#loading data
ds_raw <- read.csv2("IMBD_movie_metadata.csv",sep = ",",header = TRUE)
```
```{r}
#Data Exploration
str(ds_raw)
```
```{r}
#Data Exploration
names(ds_raw)
```

###Description of attributes

Variable Name            | Description
---------------------|----------------------------------------------
movie_title              | Title of the Movie
duration                 | Duration in minutes
director_name            | Name of the Director of the Movie
director_facebook_likes  | Number of likes of the Director on his Facebook Page
actor_1_name             | Primary actor starring in the movie
actor_1_facebook_likes   | Number of likes of the Actor_1 on his/her Facebook Page
actor_2_name             | Other actor starring in the movie
actor_2_facebook_likes   | Number of likes of the Actor_2 on his/her Facebook Page
actor_3_name             | Other actor starring in the movie
actor_3_facebook_likes   | Number of likes of the Actor_3 on his/her Facebook Page
num_user_for_reviews     | Number of users who gave a review
num_critic_for_reviews   | Number of critical reviews on imdb
num_voted_users          | Number of people who voted for the movie
cast_total_facebook_likes| Total number of facebook likes of the entire cast of the movie
movie_facebook_likes     | Number of Facebook likes in the movie page
plot_keywords            | Keywords describing the movie plot
facenumber_in_poster     | Number of the actor who featured in the movie poster
color                    | Film colorization. ‘Black and White’ or ‘Color’
genres                   | Film categorization like ‘Animation’, ‘Comedy’, ‘Romance’, ‘Horror’, ‘Sci-Fi’, ‘Action’, ‘Family’
title_year               | The year in which the movie is released (1916:2016)
language                 | English, Arabic, Chinese, French, German, Danish, Italian, Japanese etc
country                  | Country where the movie is produced
content_rating           | Content rating of the movie
aspect_ratio             | Aspect ratio the movie was made in
movie_imdb_link          | IMDB link of the movie
gross                    | Gross earnings of the movie in Dollars
budget                   | Budget of the movie in Dollars
imdb_score               | IMDB Score of the movie on IMDB

##Data Description
The dataset contain the 5043 observations of movie's metadata, with 28 different attributes such as color, duration of the movie, principal actor name, number of votes per user, and likes.
###Data types
<!--What data types are there?
What kinds of attributes are there? Categorical, numerical, character, Boolean,
continuous, etc?  -->
The dataset contain only factors that represent categorical values, integer, and numeric values. 
```{r}
#Table for know what kind of variable is define for each attribute in the data frame
sapply(ds_raw, class)
```
<!--If character, then are they also categorical or can they be unique values?-->
Character values for this dataset are always as factor, therefore, they are always categorical.
###Statistical trends
<!--If numeric, then what are the trends (mean, median,variance, etc)?-->
The summary of the raw dataset is presented below, this helps to understand generalities of each attribute. 
```{r}
#Summary of the raw data
summary(ds_raw)
```

###Definition of the target variable
The target value is the the IMBD Score that correspond to the column 26. For convinience, the target variable was moved to the end of the dataframe. 
```{r}
match("imdb_score",names(ds_raw))
ds_raw <- ds_raw[ , c(1:25,27,28,26)] #moving the target variable as the last column in the dataset 
```
```{r}
names(ds_raw)[28]
```

##Elimination of duplicate observations
There are 45 observations that are duplicated. They may be eliminated from the dataset, the target value will be moved to the end. 
```{r}
sum(duplicated(ds_raw))
ds_clean <- ds_raw[!duplicated(ds_raw), ]
```
```{r}
names(ds_clean)
```
This results in a new dataset of 4998 observations with 28 variables.

#Feature Selection

##Cardinality
  <!--What is the cardinality? Use tables and graphs to help describe information-->
  should be other type like numerical or character, the cardinality is too high
  director_name            | Factor w/ 2399 
  movie_imdb_link          | Factor w/ 4919 | This variable probably dont give me much info. 
  plot_keywords            | Factor w/ 4761
  actor_1_name             | Factor w/ 2098
  actor_2_name             | Factor w/ 3033 
  actor_3_name             | Factor w/ 3522
  #looks strange, like too many factor for what I expected. 
  genres                   | Factor w/ 914 |This one does not make much sense.
This results in a new dataset of 4998 observations with 28 variables.
From the cardinality evaluated before, it can be observed that some features may need some data cleanning.

###Genres Feature
The variable genres may not have that many factors. After analizing the data, the amount of factors is because the field contains more than one "type of genres" separated by the character pipe "|". In reality the cardinality of this variable should be 25. 

####Fixing cardinality of Genres

In order to fix the cardinality a new data frame can be created, it contains each genre class as a column. The value of the column is one if the class appeared, and zero if the class does not. The R code below implement the soluction. 
```{r}
#Code for fixing the cardinality of the genre to 25
# create a new data frame
genre_df <- as.data.frame(ds_clean[,c("genres", "imdb_score")])
# separate different genres into new columns
genre_df$Action <- sapply(1:length(genre_df$genres), function(x) if (genre_df[x,1] %like% "Action") 1 else 0)
genre_df$Adventure <- sapply(1:length(genre_df$genres), function(x) if (genre_df[x,1] %like% "Adventure") 1 else 0)
genre_df$Animation <- sapply(1:length(genre_df$genres), function(x) if (genre_df[x,1] %like% "Animation") 1 else 0)
genre_df$Biography <- sapply(1:length(genre_df$genres), function(x) if (genre_df[x,1] %like% "Biography") 1 else 0)
genre_df$Comedy <- sapply(1:length(genre_df$genres), function(x) if (genre_df[x,1] %like% "Comedy") 1 else 0)
genre_df$Crime <- sapply(1:length(genre_df$genres), function(x) if (genre_df[x,1] %like% "Crime") 1 else 0)
genre_df$Documentary <- sapply(1:length(genre_df$genres), function(x) if (genre_df[x,1] %like% "Documentary") 1 else 0)
genre_df$Drama <- sapply(1:length(genre_df$genres), function(x) if (genre_df[x,1] %like% "Drama") 1 else 0)
genre_df$Family <- sapply(1:length(genre_df$genres), function(x) if (genre_df[x,1] %like% "Family") 1 else 0)
genre_df$Fantasy <- sapply(1:length(genre_df$genres), function(x) if (genre_df[x,1] %like% "Fantasy") 1 else 0)
genre_df$`Film-Noir` <- sapply(1:length(genre_df$genres), function(x) if (genre_df[x,1] %like% "Film-Noir") 1 else 0)
genre_df$History <- sapply(1:length(genre_df$genres), function(x) if (genre_df[x,1] %like% "History") 1 else 0)
genre_df$Horror <- sapply(1:length(genre_df$genres), function(x) if (genre_df[x,1] %like% "Horror") 1 else 0)
genre_df$Musical <- sapply(1:length(genre_df$genres), function(x) if (genre_df[x,1] %like% "Musical") 1 else 0)
genre_df$Mystery <- sapply(1:length(genre_df$genres), function(x) if (genre_df[x,1] %like% "Mystery") 1 else 0)
genre_df$News <- sapply(1:length(genre_df$genres), function(x) if (genre_df[x,1] %like% "News") 1 else 0)
genre_df$Romance <- sapply(1:length(genre_df$genres), function(x) if (genre_df[x,1] %like% "Romance") 1 else 0)
genre_df$`Sci-Fi` <- sapply(1:length(genre_df$genres), function(x) if (genre_df[x,1] %like% "Sci-Fi") 1 else 0)
genre_df$Short <- sapply(1:length(genre_df$genres), function(x) if (genre_df[x,1] %like% "Short") 1 else 0)
genre_df$Sport <- sapply(1:length(genre_df$genres), function(x) if (genre_df[x,1] %like% "Sport") 1 else 0)
genre_df$Thriller <- sapply(1:length(genre_df$genres), function(x) if (genre_df[x,1] %like% "Thriller") 1 else 0)
genre_df$War <- sapply(1:length(genre_df$genres), function(x) if (genre_df[x,1] %like% "War") 1 else 0)
genre_df$Western <- sapply(1:length(genre_df$genres), function(x) if (genre_df[x,1] %like% "Western") 1 else 0)
```

**Removing the first column of the data frame of genre**
```{r}
genre_df<-subset(genre_df, select = -c(genres)) #The info is in the other columns, first one is not need it. 
```
```{r}
print(genre_df[1:5,]) 
```
```{r}
sum(is.na(genre_df)) #checking if there are NA Values
```
```{r}
# get the mean of imdb score for different genres
means <- rep(0,23) #This create a vector 
for (i in 1:23) {
  means[i] <- mean(as.numeric(genre_df$imdb_score[genre_df[i+1]==1])) #getting the mean of score for i genre
}
```
```{r}
barplot(means, main = "Average IMBD scores for different genres", xlab = "Class of Genre", ylab = "Score in IMBD")#plot of means
```
```{r}
print("Let's check the statistical values")
summary(as.numeric(ds_clean$imdb_score),na.rm=TRUE)
var(as.numeric(ds_clean$imdb_score),na.rm=TRUE)
print("Let's see how is the same statistical analisys but for the means")
summary(as.numeric(means),na.rm=TRUE)
var(as.numeric(means),na.rm=TRUE)
```
With this plot can be said that the score is pretty similar beside the genres. 
`**Removing Genres**
```{r}
#ds_clean<-subset(ds_clean, select = -c(genres)) #think about this, is this correct? 
```
**Movie Titles Removal**
The number of titles  out of 4998 observationsthat repeat in the data setset are:
```{r}
#out of 4998
sum(duplicated(ds_clean$movie_title))
```
Because few movie titles are repetead, it will be "almost" unique per each observation and therefore I do consider, it would not be a good feature to infered the IMBD score.  
**Removing movie_title**
```{r}
###REMOVING
ds_clean<-subset(ds_clean, select = -c(movie_title)) #Remove movie title as a feature
```
```{r}
names(ds_clean) #checking movie title was removed 
```

##Missing values in my data set
The dataset contains missing values, the table below summarize the missing values per feature. Because some of this values are going to be use during the classification is important to handle them. The option are either remove the observation or performe a data imputation. Also, if a feature contains several missing values it may be better to remove the feature completely, for this case any of the feature have that problem. 
```{r}
print(colSums(sapply(ds_clean, is.na)))
```
Gross and budget are the feature with more missing values, in percentage 17 percent per gross income and 1% per budget. 
```{r}
print((colSums(sapply(ds_clean, is.na)))/nrow(ds_clean))
```

###Data imputation function
As imputation method, the non available number will be replaced by the average. 
```{r}
Imputaction_function <- function(data_frame_input)
{
  for (i in (1:ncol(data_frame_input)))
    {
    if (sum(is.na(data_frame_input[,i]))!=0)
        {
          data_frame_input[(is.na(data_frame_input[,i])),i]=mean(!is.na(data_frame_input[,i]))
        }
  }
  return(data_frame_input)
}
```

###Observation elimination for high quantity of missing values
```{r}
#optional uncoment if you decide to eliminate
#ds_clean <- ds_clean[!is.na(ds_clean$gross), ]
#ds_clean <- ds_clean[!is.na(ds_clean$budget), ]
```


```{r}
#Using the function of data imputation
#print(colSums(sapply(ds_clean, is.na))) #For testing
ds_clean <- Imputaction_function(ds_clean) #calling data imputation by average
```

```{r}
print(colSums(sapply(ds_clean, is.na)))
```

##Correlation between variables related with facebook
Before performing this analisys my expectation was that all the variables related with facebook likes would be related directly with each other, however, the analisys below shows that this is not true except for actor 2 with casting and actor 3 with casting. But even in those cases the maximum correlation was 62%. With this I consider I can not eliminate this variables. 
```{r}
cor(ds_clean$director_facebook_likes,ds_clean$movie_facebook_likes)
cor(ds_clean$actor_1_facebook_likes,ds_clean$movie_facebook_likes)
cor(ds_clean$actor_2_facebook_likes,ds_clean$movie_facebook_likes)
cor(ds_clean$actor_3_facebook_likes,ds_clean$movie_facebook_likes)
cor(ds_clean$cast_total_facebook_likes,ds_clean$movie_facebook_likes)

cor(ds_clean$director_facebook_likes,ds_clean$actor_1_facebook_likes)
cor(ds_clean$director_facebook_likes,ds_clean$actor_2_facebook_likes)
cor(ds_clean$director_facebook_likes,ds_clean$actor_3_facebook_likes)
cor(ds_clean$director_facebook_likes,ds_clean$cast_total_facebook_likes)

cor(ds_clean$actor_1_facebook_likes,ds_clean$actor_1_facebook_likes)
cor(ds_clean$actor_1_facebook_likes,ds_clean$actor_2_facebook_likes)
cor(ds_clean$actor_1_facebook_likes,ds_clean$actor_3_facebook_likes)
cor(ds_clean$actor_1_facebook_likes,ds_clean$cast_total_facebook_likes)

cor(ds_clean$actor_2_facebook_likes,ds_clean$actor_1_facebook_likes)
cor(ds_clean$actor_2_facebook_likes,ds_clean$actor_3_facebook_likes)
cor(ds_clean$actor_2_facebook_likes,ds_clean$cast_total_facebook_likes)

cor(ds_clean$actor_3_facebook_likes,ds_clean$cast_total_facebook_likes)
```
```{r}
cor(ds_clean$director_facebook_likes,as.numeric(ds_clean$imdb_score))
cor(ds_clean$actor_1_facebook_likes,as.numeric(ds_clean$imdb_score))
cor(ds_clean$actor_2_facebook_likes,as.numeric(ds_clean$imdb_score))
cor(ds_clean$actor_3_facebook_likes,as.numeric(ds_clean$imdb_score))
cor(ds_clean$cast_total_facebook_likes,as.numeric(ds_clean$imdb_score))
```
```{r}
ds_clean[]
```



```{r}
names(ds_clean)
```


#Feature Selection
<!--
As part of the analysis, determine which features/attributes should be included and
which can be removed. Some variables have intuitive reasons for why they should
be included or removed. Others require appropriate justification
  The feature selection will depends on the final use of the dataset. For example, if once may want to evaluate the movie to get the imbd score, maybe some features do not affect significantly the imbd score. However, some user may be interested on sorting the data by that specific feature. In this work, I may assume that the only matter is predicting the imbd score.
  -->

#Clasification Algorithms
<!--
Apply at least 2 classification algorithms.**
▪ Discuss how you cleaned up the data
▪ Explain what data was ultimately used for each algorithm and what you had to do to prepare the data for proper use in each algorithm
-->

#Your conclusions
<!--must contain a comparison of the algorithms, the advantages and disadvantages of each, and the reason why one algorithm may have performed the best-->

#Chech you have all this sections
<!--
Introduction (5)
Data Description and Preparation (15)
Modeling Approach (15)
Analysis/Results (10)
Conclusions/Next Steps (5)
-->

